Data-Driven Modelling \& \\ Forecasting of Chaotic Systems

### 2022-01-19 projectReport/chapter4.tex
Rereading the project. On 5.4

In this chapter, we discuss mathematical results relating to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation. We then describe the \emph{single-delay dynamics} (SDD) of the system and finally state conditions on the driven system so that the SDD are conjugate (or at least semi-conjugate) to the underlying system. 
The SDD can then be used to forecast and reconstruct the underlying system.

@@

We immediately recall the figure first defined in\eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying or unknown dynamical system $(U,T)$ by making use of the measurements obtained from  the unknown system. 
See this sketched in the diagram~\ref{UV_conjugacy}

%#####################################
\begin{equation}
  % \begin{center}
    \centering
    \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
    \everypsbox{\scriptstyle}
    \begin{psmatrix}\label{UV_conjugacy}
    U & U\\%
    V & V.
    %%%
    %  \ncline{1,1}{1,2}^{T} 
    %  \ncline{1,1}{2,1} <{\phi}
    %   \ncline{2,1}{2,2}^{S}
    %  \ncline{1,2}{2,2} > {\phi}
    \end{psmatrix}
    %#####################################
    % \end{center}
\end{equation}

To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system. Two spaces are considered: input is taken from an input space $U$ and the state of a space $X$ is updated. (Previously, only the state was considered.)
We do this to mimic the true conditions. By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, it produces more general models than an autonomous system.

\begin{Definition}
  [\bf Driven, Compactly Driven Dynamical System]\label{Dfn_DDS} \rm
A driven dynamical system comprises two topological metric spaces $U$, $X$ and a continuous function  $g:U\times{X}\to{X}$ where $g(u_n, x_n)=x_{n+1}$.
If the input space $U$ is compact, we refer to the system as compactly driven. 
\end{Definition}

The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$ (a compact space) generate the dynamics on X. 
Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.

Notably, a nonautonomous dynamical system can be generated from $U$; any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps ${\{g(u_n, \cdot)\}}_{n\in\mathbb{Z}}$ contained in $X$.
Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system (alternatively, the first time a probe is inserted into the system to take an observation)

@@

\begin{Example}\label{ex_halfux} \rm
  The only solution  ${\{x_n\}}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence, say $u_n=0.5$. 
  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
\end{Example}
% \ednote{Nicely written! Thank you}

A system may also have multiple solutions, as evidenced in the example below.

\begin{Example}\label{Ex_exp} \rm 
  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$. The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $\lim_{n\to\infty}x_n=0$, $\lim_{n\to{-}\infty}x_n=1$.  
 The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
\end{Example}

As the solutions to a driven system are considered an important entity, we next identify a subspace $X_U$ of $X$ that contains all possible solutions. To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.

@@

The reachable set itself can defined independently from $g$ possessing the ESP/USP. Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.

The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$. 
Now suppose, we for some fixed input 
$\overline{u}$, we define $g_i = g(u_i,\cdot)$ for all $i\in \mathbb{Z}$, then the set of states
\begin{equation} \label{eqn_association}
X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
\end{equation}

The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
whenever $X$ is compact, the intersection is nonempty. We say $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$.

It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier. This will lead to a result (see~\cite{manjunath2013echo}) that \textit{$g$ being a topological contraction is equivalent to the existence of a exactly one entire solution.}

% \begin{Definition}
%   [\bf Topological Contraction]\label{Dfn_TopContr}\rm
%   A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subset{U}$, $X_n(\overline{u})$ is a singleton subset of $X$.  
% \end{Definition}

\begin{Remark}
  [\bf Proof of Example~\ref{Ex_exp}]\label{rem_proofEx} \rm
  For the driven system,  
  $g(u,x)=x^2$ the evolution does not depend on the input at all. Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$. A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$. Or in other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$ 
\end{Remark}

Thus far, it has been demonstrated that a system may have one or more solutions; one may ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness. 
Should the driven system be compact, existence follows immediately, as shown in the following result.

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\vspace{-6mm}
{\bf Proof.} May be found in~\cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$. To refine the scenario, we consider only systems with unique solutions. 
\section{Unique Solution Property}

\begin{Definition}
  [\bf Unique Solution Property]\label{Dfn_usp}\rm
  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, with $\Psi({\overline{u}})$ denoting the unique solution.
\end{Definition}

One of the first results obtained after defining the USP is that every solution will attract different initial conditions towards the component parts of the solution.
If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}. The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to~\cite{Manju_ESP, esann2012ids} for further reading.

% Paraphrased.
%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).

In the majority of Reservoir Computing (RC) literature- a machine learning approach- where the input is mapped onto a different but higher dimensional through a system called the reservoir and a readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.

% \ednote{How should I explain/qualify RC better here? It's mentioned without really explaining. M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir". B: Thanks!}, 
Concepts like the ESP,  fading memory~\cite{boyd1985fading} are some of these notions.  
If $g$ possesses the ESP (equivalent in our context to the USP), the we are guaranteed that the whole of an input's left-infinite history will exactly determine the current system state; i.e. there is only ever the possibility of a single reachable state at a given instance of time~\cite{jaeger2001echo,Manju_2020}.
% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described. A more detailed presentation may be found in~\cite{jaeger2001echo}.

% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
Beyond forgetting the past states the ESP is also often discussed in terms of a stability property that is nearby inputs create close by responses (solutions), a concept called input-related stability~\cite{manjunath2020stability} and, furthermore, plays a key role in the robustness of recurrent neural networks (RNN), a component of the implementation discussed in~\ref{ch5}.

@@

Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ is then the notation for all these left-infinite sequences in $U$. 
Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$ with $v \in U$ being the specific input value at time $n$. 
The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$

The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below.

\begin{equation}  \label{Scomm_h}
  %    \[ 
      \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
   \everypsbox{\scriptstyle}
   \begin{psmatrix}
   \overleftarrow{U} & \overleftarrow{U}\\%
   X_U & X_U.
   %%%
  %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
  %  \ncline{1,2}{2,2} > {h}
  %  \ncline{2,1}{2,2}^{g(v,\cdot)}
   \end{psmatrix}
  % \]
  \end{equation}

@@

But does such a function $h$ for the above driven system $g$ above exist? Whenever $g$ has the USP and $\Psi(u)={\{x_n\}}_{n\in\mathbb{Z}}$ it follows that $h$, defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$, will satisfy the semi-conjugacy in the graph above~\ref{Scomm_h}.
Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP~\cite[Lemma 5]{Manju_Nonlinearity}.
Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example~\ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.

Re-sketching the commutativity diagram~\ref{Scomm_h} above by replacing $X_U$ by its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram below. In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping} that is defined next.

\begin{Definition}
  [\bf Causal Mapping]\label{Def_CausMap}
  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
\end{Definition}

\begin{equation} \label{SCausal_H}
    %    \[ 
        \psset{arrows=->, arrowinset=0.25, linewidth=0.6pt, nodesep=3pt, labelsep=2pt, rowsep=0.7cm, colsep = 1.1cm, shortput =tablr}
     \everypsbox{\scriptstyle}
     \begin{psmatrix}
     \overleftarrow{U} & \overleftarrow{U}\\%
     \overleftarrow{X}_U & \overleftarrow{X}_U.
     %%%
    %  \ncline{1,1}{1,2}^{\sigma_v} \ncline{1,1}{2,1} <{h}
    %  \ncline{1,2}{2,2} > {h}
    %  \ncline{2,1}{2,2}^{\tilde{g}_v}
     \end{psmatrix}
    % \]
  \end{equation}

 \begin{Theorem}
  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
\end{Theorem}
\vspace{-6mm}
{\bf Proof.}  See~\cite[Th.3]{manjunath2013echo}.

The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to being surjective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$. 
The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
and we refer to the map $H$ as a \emph{causal embedding}. 
%\ednote{B: edited the sentence here}

When $g$ has the USP, the diagram below (taken from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$. The mapping $h:\overleftarrow{U}\to{X_U}$ is also considered an observable as mentioned in the introduction of Chapter~\ref{ch3}).

\begin{figure}[ht]
  \includegraphics[scale=0.4]{Graphs/_actionofh_H.eps}
  \centering
  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
 \label{fig:actionh_H}
\end{figure}

The function $H$ contains a countably infinite number of coordinate (or factor maps) since its codomain is $\overleftarrow{X}_U$.  
As it is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning, we wish to restrict the inputs we consider to a subspace of $\overleftarrow{U}$. 
Particularly, we restrict our attention to the  so-called inverse-limit space of a dynamical system and hope to embed it within some finite product of $X$.

@@

The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$. 
Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition:

>>>>

In this chapter, we discuss mathematical results relating to the mapping of temporal data obtained from a discrete-dynamical system onto a different space through the notion of a driven dynamical system. 
We also consider the conditions a driven system should possess to avoid adding distortion to its state-space representation. 
We then describe the \emph{single-delay dynamics} (SDD) of the system, and finally state conditions on the driven system for the SDD to be conjugate (or at least semi-conjugate) to the underlying system. 
The SDD can then be used to forecast and reconstruct the underlying system.

@@

We immediately recall the figure first defined in\eqref{eqn_conjugacy}  and remind ourselves of our ultimate goal to learn a system that is conjugate to an underlying/unknown dynamical system $(U,T)$ by making use of measurements obtained from  this unknown system. 
We again sketch the commutativity diagram~\ref{eqn_conjugacy}.
\begin{equation*}\label{UV_conjugacy}
  \begin{tikzcd}
    U && U \\
    V && V \\
    \arrow["T", from=1-1, to=1-3]
    \arrow["\phi", from=1-3, to=2-3]
    \arrow["\phi"', from=1-1, to=2-1]
    \arrow["S"', from=2-1, to=2-3]
  \end{tikzcd}
  \end{equation*}

To find a function $\phi$, the need now arises to consider a so-called driven dynamical system, a special case of a nonautonomous dynamical system. 
Two spaces are considered: input is taken from an input space $U$, and the state of a space $X$ is updated. (Previously, only the state was considered.)
We do this to mimic the true conditions. By accounting for an extrogenous input sequence in $U$ which can also influence the  system-state at a specific time-step, models more general than an autonomous system are produced.

\begin{Definition}
  [\bf Driven and Compactly Driven Dynamical System]\label{Dfn_DDS} \rm
A driven dynamical system comprises two topological metric spaces $U$ and $X$, as well as  a continuous function  ${g:U\times{X}\to{X}}$ where ${g(u_n, x_n)=x_{n+1}}$.
If the input space $U$ is compact, we refer to the system as compactly driven. 
\end{Definition}

The update equation $x_{n+1} = g(u_n,x_n)$ for $n \in\mathbb{Z}$, input $u_n$ from $U$ and state $x_n$ belonging to $X$,a compact space, generates the dynamics on $X$. 
Abbreviated, we refer only to the \textit{driven system $g$}, where all other entities are understood implicitly.

Notably, a nonautonomous dynamical system may be generated from $U$. Any input $\overline{u}$, a bi-infinite sequence from $U$, gives rise to the sequence of self-maps ${\{g(u_n, \cdot)\}}_{n\in\mathbb{Z}}$ contained in $X$.
Physically, one may think of this bi-infinite sequence as referring to a system that has been running for an extended period at the time of the first measurement taken from the system. Alternatively, one may conceive of the scenario where this is the first time a probe is inserted into a system to take an observation.

@@

\begin{Example}\label{ex_halfux}\rm
  The only solution  ${\{x_n\}}_{n\in\mathbb{Z}}$ to the driven system  $g(u,x)=\frac{ux}{2}$, where $X=[0,1]$, $U=[0,1]$,  is the zero solution $x_n\equiv0$.
  To see this, consider any $x_n=a\in[0,1]$ where $a\neq{0}$.  Let $\overline{u}\in{U}$ be an non-zero constant sequence, say $u_n=0.5$. 
  The driven system may be rewritten as $x_{n-1}=\frac{2x_n}{u_{n-1}}=4x_n$ and the  iterates of $x_n$ in backward time will increase by a factor of 4 at each timestep. 
  Thus for some $m\leq{n}$,  $1<x_m$ i.e. $x_m\notin{X}$. So ${\{x_n\}}_{n\in\mathbb{Z}}$ is not a solution and it follows that the only possible solution is the zero solution.
\end{Example}
% \ednote{Nicely written! Thank you}

A system may also have multiple solutions, as evidenced in Example~\ref{Ex_exp}.

\begin{Example}\label{Ex_exp}\rm 
  Consider the driven system $g(u,x)=x^2$ for $X=[0,1]$, $U=\mathbb{R}$. The system has an uncountable number of solutions as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $\lim_{n\to\infty}x_n=0$, $\lim_{n\to{-}\infty}x_n=1$.  
 The proof is deferred to immediately after the next paragraph (see~\ref{rem_proofEx}).
\end{Example}

Solutions to a driven system are considered an important entity, so we next identify a subspace $X_U$ of $X$ that contains all possible entire solutions. 
% To realize such a subspace of a driven system $g$, the concept of a reachable set is defined.

@@

\ednote{Moved to after defn of USP/ESP: The reachable set itself can defined independently from $g$ possessing the ESP/USP. } 
Note that $x\in{X_n(\overline{u})}$ if and only $g$ has a solution $\{x_k\}$ for $x_n=x$ and input $\overline{u}$, a result proved in nonautonomous dynamical systems literature~\cite{manjunath2014dynamics, manjunath2013echo}.

The set $X_n(\overline{u})$ is precisely the set of points $x$ through which some entire solution $\Psi$ passes at time $n$. We can give an alternate formulation of the set $X_n(\overline{u})$: 
Suppose for some fixed input 
$\overline{u}$, we define $g_i = g(u_i,\cdot)$ where $i\in \mathbb{Z}$, then the set of states
\begin{equation} \label{eqn_association}
X_{n,i}(\overline{u}) := g_{n-1} \circ \cdots g_{i+1} \circ g_i(X).
\end{equation}

The set $X_{n,i}(\overline{u})$ being the image of a finite composition of
continuous maps is compact whenever $X$ is compact. Also $X_{n,i}(\overline{u})$ is
nonempty.  Further, $X_{n,i}(\overline{u}) \supset X_{n,i-1}$. Hence $\bigcap_{i<n}
X_{n,i}(\overline{u})$ is a nested intersection of closed nonempty subsets, and
whenever $X$ is compact, the intersection is nonempty.

\begin{Definition}
  [\bf Topological Contraction]\label{Dfn_topContr}\rm
  The driven system $g$ is a topological contraction if $\bigcap_{i<n}X_{n,i}(\overline{u})$ is a singleton subset of $X$ for each $\overline{u}$. 
\end{Definition}

It turns out that $\bigcap_{i<n}X_{n,i}(\overline{u})$ is identically equal to $X_n((\overline{u})$ that we had defined earlier. This will lead to a result (see~\cite{manjunath2013echo}) that 
\textit{$g$ being a topological contraction is equivalent to the existence of exactly one entire solution.}

\begin{Remark}
  [\bf Proof of Example~\ref{Ex_exp}]\label{rem_proofEx} \rm
  For the driven system,  
  $g(u,x)=x^2$ the evolution does not depend on the input at all. Hence $g(u,x)$ can be written as a single map $f(x)=x^2$ which is a homeomorphism on $[0,1]$. 
  A left-infinite orbit is defined and it converges to $1$ while the right-infinite orbit converges to $0$. 
  In other words the iterates of $f^{-1}$ converges to $1$ while the iterates of $f$ converge to $0$ 
\end{Remark}

Thus far, it has been demonstrated that a system can have a single or multiple solutions. One may therefore reasonably ask if a driven system always has a solution and, if so, whether it satisfies specific properties such as uniqueness. 
Should the driven system be compact, existence follows immediately, as shown in the following result.

\begin{Theorem}\label{Thm_CompactExistence}
 Let $g$ be a driven system.  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\vspace{-6mm}
{\bf Proof.} May be found in~\cite{kloeden2011nonautonomous, manjunath2014dynamics, manjunath2013echo}
% One may easily construct many systems with trivial solution-sets, such as $g(u,x)=x$ which has only the constant solution $x$ and so for $U=[-1,1]$, the system would have no solution if $|x|>1$. To refine the scenario, we consider only systems with unique solutions.

\section{Unique Solution Property}

\begin{Definition}
  [\bf Unique Solution Property]\label{Dfn_usp}\rm
  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$, where $\Psi({\overline{u}})$ denotes the unique solution.
\end{Definition}

The USP is equivalently called  the echo state property (ESP)~\cite{Manju_ESP}, and systems that possess the ESP have other important and useful properties for information processing~\cite{grigoryeva2018echo, grigoryeva2019differentiable, manjunath2020stability} .

One of the first results obtained after defining the USP refers to the fact that every solution will attract different initial conditions towards the component parts of the solution.
If $g$ has the USP, then any solution to $g$ is also a uniform attractor in nonautonomous dynamical systems literature~\cite{Manju_Nonlinearity}. The discussion on nonautonomous attractors is beyond the scope of this project, and we refer the reader to~\cite{Manju_ESP, esann2012ids} for further reading.

Secondly, it is worth noting that the reachable set~\ref{Dfn_ReachableSet} can be defined independently from $g$ possessing the USP. \ednote{hmm. does it fit here}
% Paraphrased.
%Additionally, $g$ having the USP is in our context also equivalent to $g$ being a topological contraction \cite{manjunath2021universal}, or alternatively also that $g$ exhibits the Echo State Property (ESP).

In the majority of Reservoir Computing (RC) literature -- a machine learning approach -- where the input is also mapped onto a different, but higher dimensional, space through a system called the reservoir. A readout that measures the state of the reservoir is trained, a notion of forgetting the states of the reservoir asymptotically is often used.   \ednote{this sentence falls out of the blue}

% \ednote{How should I explain/qualify RC better here? It's mentioned without really explaining. M: Have put it in one sentence; just to let you know people use optical devices to map signal into a ``reservoir". B: Thanks!}, 
Concepts like the echo state property (ESP) and fading memory~\cite{boyd1985fading} are some of these notions.  
If $g$ possesses the ESP (equivalent in our context to the USP), then we are guaranteed that the whole of an input's left-infinite history will determine exactly the current system state; i.e. there is only ever the possibility of a single reachable state at a given instance of time~\cite{jaeger2001echo,Manju_2020}.
% Defined formally, the ESP is intrinsically linked to an input (see \cite{Manju_ESP}) Unfortunately, definitions and results concerning the ESP do not have much practical benefit since a relationship between the network's response and the temporo-statistical properties of the input are not directly described. A more detailed presentation may be found in~\cite{jaeger2001echo}.

% We do however consider it important to mention that the ESP is also often discussed in terms of a stability property that holds for all inputs of a system~\cite{manjunath2020stability} and, furthermore, plays a key role in the design and training of recurrent neural networks(RNN) within the field of RC \cite{Manju_ESP}.
Beyond forgetting the past states, the ESP is also often discussed in terms of a stability property termed input-related stability where nearby inputs create nearby responses (solutions)~\cite{manjunath2020stability}. Furthermore, it plays a key role in the robustness of recurrent neural networks (RNN), a component of the implementation discussed in~\ref{ch5}.

@@

Having already defined the reachable set $X_U$ as the collection of all elements of all entire solutions, we pause briefly in order to fix additional notation.
Defining $\cev{u}^{n}:=(\ldots,u_{n-2} ,u_{n-1})$ as the left-infinite subsequence of an input up until time $n$, $\overleftarrow{U}$ then denotes all such left-infinite sequences in $U$. 
Moreover, $\cev{u}^{n}v:=(\ldots,u_{n-2} ,u_{n-1}, v)$ will symbolise the input up to time $n$, with $v \in U$ being the specific input value at time $n$. 
The introduction of a new input at time $n$ can be described by the mapping $\sigma_v:   \cev{u}^{n} \mapsto \cev{u}^{n}v$. 
The right-shift map, $r\overleftarrow{U}\to\overleftarrow{U}$, of an input sequence is defined $r: (\cdots, u_{-2},u_{-1}) \mapsto(\cdots, u_{-3},u_{-2})$

The question now assumes the form: Can we establish a semi-conjugacy for the driven system $g$ as presented below?

  \begin{equation}\label{Scomm_h}
    \begin{tikzcd}
	{\overleftarrow{U}} && {\overleftarrow{U}} \\
	{X_U} && {X_U} \\
	\arrow["{\sigma_v}", from=1-1, to=1-3]
	\arrow["h"', from=1-3, to=2-3]
	\arrow["h", from=1-1, to=2-1]
	\arrow["{g(v,\cdot)}"', from=2-1, to=2-3]
\end{tikzcd}
\end{equation}

@@

But does such a function $h$ for the above driven system $g$ above exist? Whenever $g$ has the USP and $\Psi(u)={\{x_n\}}_{n\in\mathbb{Z}}$, it follows that $h$ defined by  $h(\cev{u}_n):=g(u_n,x_{n-1})=x_n$ will satisfy the semi-conjugacy in the graph above~\ref{Scomm_h}.
Regrettably, such a continuous mapping $h$ is not guaranteed to exist when $g$ does not have the USP.

\begin{Theorem}
  A compactly driven system $g$ induces a universal semi-conjugacy $h:\overleftarrow{U}\to{X_U}$ if and only if $g$ has the USP. 
\end{Theorem}
\vspace{-6mm}
{\bf Proof.} See the proof of Lemma 5 provided in~\cite{Manju_Nonlinearity}

Note that even when $h$ does exist, we are not guaranteed its injectivity. In example~\ref{ex_halfux} we observe that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.

Re-sketching the commutativity diagram~\ref{Scomm_h} above by replacing $X_U$ with its left-infinite sequence space $\overleftarrow{X}_U$, we obtain the diagram~\ref{SCausal_H}. 
In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.

\begin{Definition}
  [\bf Causal Mapping]\label{Def_CausMap}
  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma_v \circ{H}\] holds for all $v \in U$ where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$ is called a causal mapping.
\end{Definition}

  \begin{equation}\label{SCausal_H}
    \begin{tikzcd}
    \overleftarrow{U} && \overleftarrow{U}\\
     \overleftarrow{X}_U && \overleftarrow{X}_U.
     \\
    \arrow["{\sigma_v}", from=1-1, to=1-3]
	\arrow["h"', from=1-3, to=2-3]
	\arrow["h", from=1-1, to=2-1]
	\arrow["{\tilde{g_v}}"', from=2-1, to=2-3]
\end{tikzcd}
\end{equation}

 \begin{Theorem}
  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
\end{Theorem}
\vspace{-6mm}
{\bf Proof.}  See~\cite[Theorem.3]{manjunath2013echo}.

The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: 
If the causal mapping $H:\overleftarrow{U}{\to}{\overleftarrow{X}_U}$ is injective (in addition to its surjectivity), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$. 
The continuity of $H^{-1}$ follows from the fact that $H$ is itself a continuous and surjective mapping of a compact space $\overleftarrow{U}$ in a Hausdorff space.
and we refer to the map $H$ as a \emph{causal embedding}. (Compare this with Definition~\ref{defn_causallyEmbed}.) 
%\ednote{B: edited the sentence here}

When $g$ has the USP, the diagram below (taken from~\cite{Manju_Nonlinearity}) illustrates the operation of the mappings $h$ and $H$. 
The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as mentioned in the introduction of Chapter~\ref{ch3}).

\begin{figure}[ht]
  \includegraphics[width=0.6\linewidth]{Graphs/_actionofh_H.eps}
  \centering
  \captionof{figure}{Ladder-like behaviour of $h$ and $H$ to illustrate the causal mapping. (Figure reproduced from~\cite{Manju_Nonlinearity}) }
\label{fig:actionh_H}
\end{figure}

The function $H$ contains a countably infinite number of coordinate (or \textit{factor maps}) since its codomain is $\overleftarrow{X}_U$.  
It is not practically feasible to consider a left-infinite sequence in $X_U$ (in other words an element of  $\overleftarrow{X}_U$) for learning. We wish to restrict the inputs to a subspace of $\overleftarrow{U}$. 
Particularly, we restrict our attention to the  so-called inverse-limit space (see Definition~\ref{Dfn_InverseL}) of a dynamical system and hope to embed it within a finite product of $X$.

@@

The left-infinite orbit space is significant because a great amount of information about the original system can be gleaned from its topological structure~\cite{ingram2011inverse,Manju_IEEE}. 
The formal topological structure for the inverse-limit space of a chaotic dynamical system is complicated, and further discussion is relegated to~\cite{kennedy2008inverse_limit, ingram2011inverse}.
Broadly, we may conceive of the inverse-limit system of a dynamical system as a subspace of an infinite-dimensional space where each point in the inverse-limit space corresponds to a backward orbit of the map $T$. 
Following~\cite{manjunath2021universal, ingram2011inverse} we provide a definition.
\begin{Definition}
  [\bf Inverse-Limit Space, Inverse-Limit System]\label{Dfn_InverseL}\rm
 The subset $\widehat{U}_T\subset \overleftarrow{U}$ defined by 
 \begin{equation}
  \widehat{U}_T: = \{ (\ldots,u_{-2},u_{-1}): Tu_{i} = u_{i+1}\}  
 \end{equation}
 is called the inverse-limit space of $(U,T)$.
 The self-map $\widehat{T}$  induced by $T$ on $\widehat{U}_T$  is defined by 
 \begin{equation}
  \widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))  
 \end{equation}
 and the resulting dynamical system $(\widehat{U_T}, \widehat{T})$ is called the inverse-limit system of $(U,T)$.
\end{Definition}
